---
title: "Model Predictive Control of Building Dynamics"
output: html_notebook
author: Edward J. Xu
date: Sep 1, 2019
---

```{r, include=FALSE}
rm(list = ls())
```

## Initialization

```{r}
library(MatrixTSA)
library(tibble)
library(tidyverse)
options(tibble.width = Inf)

source("./src/asHours.R")
source("./src/asP.R")
source("./src/prbs.R")
source("./src/sim_building.R")
```

If you have a working GLPK solver on your system and want to implement a
computationally more optimized controller use the following

- install.packages("curl")
- install.packages("devtools")
- library(devtools)
- install_version("slam",version="0.1-40", repos = "http://cran.us.r-project.org")

Now install the Rglpk which is used for solving linear programs

- install.packages("Rglpk")
- library(Rglpk)

## 1, Get Data

```{r, message=FALSE}
ti <- read_csv(
  "~/GitHub/MatrixTSA/data/data_mpc.csv",
  skip = 1,
  col_names = c("timedate", "Y1", "Y2", "Ta", "Gv", "Ph1", "Ph2")
)
ti <- ti %>%
  mutate(timedate = asP(.$timedate)) %>%
  mutate(t = asHours(.$timedate - .$timedate[1]))
```

# 2, Model using CTSM

```{r}
ti_est <- tibble(
  name = c(
    "T1a0", "T1m0", "C1a", "C1m", "R1a",
    "R1m", "A1w", "p1", "p1a", "p1m", "e11"
  ),
  init = c(25, 25, 6, 12, 10, 1, 1, 0.5, 1, 1, -1),
  lb = c(0, 0, 1E-5, 1, 1, 1E-10, 1E-10, 0, -30, -30, -50),
  up = c(35, 35, 20, 50, 80, 10, 10, 1, 10, 10, 10)
)
li_mod <- list()
li_mod[[1]] <- set_mod_ctsm(
  c_expr_sys = c(
    d(T1a) ~ (1 / (C1a * R1m) * (T1m - T1a) + 1 / (C1a * R1a) * (Ta - T1a) + 1 / C1a * Ph1 + p1 * A1w / C1a * Gv) * d(t) + exp(p1a) / C1a * d(w1a),
    d(T1m) ~ (1 / (C1m * R1m) * (T1a - T1m) + (1 - p1) * A1w / C1m * Gv) * d(t) + exp(p1m) / C1m * d(w1m)
  ),
  expr_obs = Y1 ~ T1a,
  expr_var = Y1 ~ exp(e11),
  c_input = c("Ta", "Ph1", "Gv"),
  ti_est = ti_est
)
```

# 3, Estimate the Model

```{r}
fit1 <- est_mod_ctsm(mod = li_mod[[1]], ti_data = ti)
```

# 4, Get State Space Model

```{r}
li_mat_ss <- trans_ctsm_ss(fit1)
li_mat_ss_d <- trans_mat_ss(li_mat_ss = li_mat_ss, ti_data = ti)
```

# 5, Simulate the Control Process
```{r}
result <- sim_building(li_mat_ss, li_mat_ss_d, ti)
Tall <- result$Tall
Ymin <- result$Ymin
Ymax <- result$Ymax
Price <- result$Price
Tmax <- result$Tmax
u <- result$u
```

# 6, Plot the Result
```{r}
plot(Tall[1, ], type = "l", ylim = c(Ymin - 0.5, Ymax + 0.5), ylab = "Air Temperature")
lines(Tall[2, ], type = "s", col = alpha("blue", 0.4))
lines(u * 0.5 + Ymin + 0.5, type = "s", col = "red")
lines((Price - 1) * 0.6 + Ymin + 1.5, col = "green", type = "s")
lines(c(Ymax, Ymax) ~ c(0, Tmax), lty = 2)
lines(c(Ymin, Ymin) ~ c(0, Tmax), lty = 2)
legend(70, max(Ymax + 0.6), c("Air Temperature", "Thermal Mass Temperature", "Heating", "Price"), col = c("Black", alpha("blue", 0.4), "red", "green"), lty = 1, bty = "n")
```

